---
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(leaflet)
library(leafem)
```


<link rel="stylesheet" href="https://unpkg.com/leaflet@1.7.1/dist/leaflet.css" integrity="sha512-xodZBNTC5n17Xt2atTPuE1HxjVMSvLVW9ocqUKLsCC5CXdbqCmblAshOMAS6/keqq/sMZMZ19scR4PsZChSR7A==" crossorigin="" />

<script src="https://unpkg.com/leaflet@1.7.1/dist/leaflet.js" integrity="sha512-XQoYMqMTK8LvdxXYG3nZ448hOEQiglfqkJs1NOQV44cWnUrBc8PkAOcXy20w0vlaXaVUearIOBhiXZ5V3ynxwA==" crossorigin=""></script>

<script src="https://d3js.org/d3.v7.min.js"></script>

<script src="https://unpkg.com/georaster"></script>

<script src="https://unpkg.com/georaster-layer-for-leaflet/dist/georaster-layer-for-leaflet.min.js"></script>

```{r, warning = FALSE, echo = F}
leaflet(width = 0, height = 0)
```


<style>
        /* Overall map style */
        #map {position: absolute; top: 0; bottom: 0; left: 0; right: 0;}

        .legend {
            padding: 6px 8px;
            font: 14px Arial, Helvetica, sans-serif;
            background: white;
            background: rgba(255, 255, 255, 0.8);
            /*box-shadow: 0 0 15px rgba(0, 0, 0, 0.2);*/
            /*border-radius: 5px;*/
            line-height: 24px;
            color: #555;
        }

        .legend h4 {
            text-align: left;
            font-size: 16px;
            margin: 2px 12px 8px;
            color: #777;
        }

        .legend span {
            position: relative;
            margin: 0 0px 0 8px;
            bottom: 3px;
        }

        .legend i {
            width: 18px;
            height: 18px;
            float: left;
            margin: 0 8px 0 14px;
            opacity: 1;
        }

        .legend i.icon {
            background-size: 18px;
            background-color: rgba(255, 255, 255, 1);
        }
	</style>


<div id="map" style="width: 100%; height: 480px; position: relative;"></div>

<script>
  // Load base tiles ----

  // OpenStreetMap Tiles
    // Dark Matter Tiles
  var DarkMatter = L.tileLayer('https://cartodb-basemaps-{s}.global.ssl.fastly.net/dark_all/{z}/{x}/{y}.png', {
    attribution: '&copy; <a href="https://carto.com/attributions">CARTO</a>'
  });

  // Initiate leaflet ----
  // This needs to be done before loading the Georasters
  var map = L.map('map', { 
      center: [56.25, 12], // Centered on Denmark
      zoom: 7, // Worked well
      layers: [DarkMatter] // Default layers
  });

  // Set URL to cog
  var random_rast_dk_url = "https://s3.eu-central-1.amazonaws.com/sustainscapes.data/CS1Int.tif"

  // Load georaster layer and add to map
  parseGeoraster(random_rast_dk_url).then(georaster => {
      var random_rast_dk = new GeoRasterLayer({
          georaster,
          opacity: 0.75,
          resolution: 128, // 128 seems to be a good compromise
          pixelValuesToColorFn: function(pixelValues){
            var pixelValue = pixelValues[0]; 
            // Define the color scale range
            var colorScale = d3.scaleSequential(d3.interpolateViridis)
                               .domain([1, 62]); // Adjust the domain to cover the range of your pixel values

            // Return color based on pixel value
            return colorScale(pixelValue);
        }
    });       
    random_rast_dk.addTo(map);
});

// Define legend
var legend = L.control({ position: "bottomright" });
legend.onAdd = function(map) {
    var div = L.DomUtil.create("div", "legend");
    
    div.innerHTML += "<h4>Projected landuse</h4>";
    div.innerHTML += '<div class="legend-gradient" id="legend-gradient"></div>';
    div.innerHTML += '<span class="legend-min">1</span><span class="legend-max">62</span>'; // Adjust the min and max values

    return div;
};

// Add legend
legend.addTo(map);



</script>

