---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(terra)
library(tidyterra)
library(ggplot2)
```

# Vilhelmsborg

<!-- badges: start -->
<!-- badges: end -->

The goal of Vilhelmsborg is to find around 56.066330, 10.185723. a 500 meter radius and 1000 meter radius, and place 1 plot per hectare in the protected area ocuring withing

```{r}
Circle <- terra::vect(data.frame(lon = 10.185723, lat = 56.066330), geom=c("lon", "lat"), crs="+proj=longlat +datum=WGS84") |>  
  terra::project("EPSG:25832") |> 
  terra::buffer(1000)

terra::writeVector(Circle, "CircleVilhelm.shp", overwrite = TRUE)


Circle_1500 <- terra::vect(data.frame(lon = 10.185723, lat = 56.066330), geom=c("lon", "lat"), crs="+proj=longlat +datum=WGS84") |>  
  terra::project("EPSG:25832") |> 
  terra::buffer(1500)

terra::writeVector(Circle_1500, "Circle_1500_Vilhelm.shp", overwrite = TRUE)

Circle_2000 <- terra::vect(data.frame(lon = 10.185723, lat = 56.066330), geom=c("lon", "lat"), crs="+proj=longlat +datum=WGS84") |>  
  terra::project("EPSG:25832") |> 
  terra::buffer(2000)

terra::writeVector(Circle_2000, "Circle_2000_Vilhelm.shp", overwrite = TRUE)

```

This has `r round(expanse(Circle, unit = "ha"), 2)` hectares.

The steps to go in this project is:

* Find out from Paragraph 3 and Natura 2000 protected areas within this area
* For each protected area within this calculate the area in Hectares
* We will test to round up or down, for each hectare of protected area, we need one plot
* This plot should be at least 10 meters from the border and tentatively 50 meters from each other
* Extra: A interactive map will be made of this

## Read in Natura 2000 and paragraph 3

From the biodiversity council read in the layers

```{r ReadLayers, cache = T}

Paragraph3 <-  terra::vect("O:/Nat_BDR-data/Arealanalyse/RAW/BES_NATURTYPER_SHAPE") |> 
  terra::project("EPSG:25832") |> terra::intersect(Circle)

Paragraph3$Status <- iconv(Paragraph3$Status, from="Windows-1252", to="UTF-8")

Paragraph3$CVR_navn <- iconv(Paragraph3$CVR_navn, from="Windows-1252", to="UTF-8")

Paragraph3$Natyp_navn <- iconv(Paragraph3$Natyp_navn, from="Windows-1252", to="UTF-8")


nfiles <- nrow(Paragraph3)

terra::writeVector(Paragraph3, "Paragraph3Vilhelm.shp", overwrite = TRUE)



Paragraph3_1500 <-  terra::vect("O:/Nat_BDR-data/Arealanalyse/RAW/BES_NATURTYPER_SHAPE") |> 
  terra::project("EPSG:25832") |> terra::intersect(Circle_1500)

Paragraph3_1500$Status <- iconv(Paragraph3_1500$Status, from="Windows-1252", to="UTF-8")

Paragraph3_1500$CVR_navn <- iconv(Paragraph3_1500$CVR_navn, from="Windows-1252", to="UTF-8")

Paragraph3_1500$Natyp_navn <- iconv(Paragraph3_1500$Natyp_navn, from="Windows-1252", to="UTF-8")


nfiles_1500 <- nrow(Paragraph3_1500)

terra::writeVector(Paragraph3_1500, "Paragraph3Vilhelm_1500.shp", overwrite = TRUE)


Paragraph3_2000 <-  terra::vect("O:/Nat_BDR-data/Arealanalyse/RAW/BES_NATURTYPER_SHAPE") |> 
  terra::project("EPSG:25832") |> terra::intersect(Circle_2000)

Paragraph3_2000$Status <- iconv(Paragraph3_2000$Status, from="Windows-1252", to="UTF-8")

Paragraph3_2000$CVR_navn <- iconv(Paragraph3_2000$CVR_navn, from="Windows-1252", to="UTF-8")

Paragraph3_2000$Natyp_navn <- iconv(Paragraph3_2000$Natyp_navn, from="Windows-1252", to="UTF-8")


nfiles_2000 <- nrow(Paragraph3_2000)

Ownership <- terra::vect("O:/Nat_BDR-data/Arealanalyse/RAW/Archive/ejerskab_20220609.gpkg")
Ownership$Ownership <- ifelse(Ownership$ejerforhold_dni %in% c("fond", "andet"), "Privat",
                              ifelse(Ownership$ejerforhold_dni %in% c("naturstyrelsen", "forsvaret","landbrugsstyrelsen", "kystdirektoratet"), "Statslig", "Kommunal"))

Ownership <- Ownership["Ownership"]

Ownership <- Ownership |> 
  terra::project("EPSG:25832") |> 
  terra::crop(Circle_2000) |> 
  terra::aggregate("Ownership") |> 
  terra::intersect(Paragraph3_2000)

terra::writeVector(Ownership, "Ownership.shp", overwrite = TRUE)

terra::writeVector(Paragraph3_2000, "Paragraph3Vilhelm_2000.shp", overwrite = TRUE)
```

There are `r nfiles` paragraph 3 areas within the Vilhemsborg area, in the following figure, paragraph 3 areas are in green and the 1000 km radius circle inside in red

```{r P3plot, cache = T}
plot(Circle, col = "red")
plot(Paragraph3, col = "green", add = T)

```

## More distances

For paragraph 3 we made 2 more distances, 1500 and 2000 meters, with `r nfiles_1500` polygons at 1500 meters and `r nfiles_2000` at 2000 meters of distance.


We can see the representation of this in the following plot:


```{r}
Circle_1000 <- terra::vect("CircleVilhelm.shp")
Circle_1000$Distance <- 1000

Circle_1500 <- terra::vect("Circle_1500_Vilhelm.shp")
Circle_1500$Distance <- 1500


Circle_2000 <- terra::vect("Circle_2000_Vilhelm.shp")
Circle_2000$Distance <- 2000

Circles <- list(Circle_1000, Circle_1500, Circle_2000) |> purrr::reduce(rbind)

Circles$Distance <- as.character(Circles$Distance)

Sites <- terra::vect("Paragraph3Vilhelm_2000.shp")
Sites$Natyp_navn <- stringr::str_replace_all(Sites$Natyp_navn, "<f8>", "ø")


ggplot() + geom_spatvector(data = Circles, aes(color = Distance), alpha = 0) + scale_color_viridis_d() +
  geom_spatvector(data = Sites, aes(fill = Natyp_navn))  + theme_minimal()


```

