---
title: "Class 3 Models"
author: "Derek Corcoran"
date: "`r format(Sys.time(), '%d/%m, %Y')`"
output:
  ioslides_presentation:
    widescreen: true
    incremental: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, cache = FALSE, tidy = TRUE, tidy.opts = list(width.cutoff = 50))
library(tidyverse)
library(broom)
library(kableExtra)
library(knitr)
options("kableExtra.html.bsTable" = T)
```

# Models

## What is a model?

* A model is a simplified version of reality that allows us to make inferences or predictions about a population
* A model is an adequate summary of reality
* A model is a simplification or approximation to reality and therefore will not reflect all of reality (Burnham and Anderson)
* All models are wrong, some are useful (George Box)

## Let's see an example

* ¿How much $CO_2$ do plants capture?

```{r, echo = FALSE, cache = TRUE}
data("CO2")
knitr::kable(CO2, row.names = FALSE) %>% kable_styling(bootstrap_options = c("striped")) %>% scroll_box(height = "800px", width = "650px")
```

## ¿Is it the subspecies?

```{r}
ggplot(CO2, aes(x = Type, y=uptake)) + geom_boxplot(aes(fill = Type), notch = TRUE) + theme_classic()
```

## ¿Is it the treatment?

```{r}
ggplot(CO2, aes(x = Treatment, y=uptake)) + geom_boxplot(aes(fill = Treatment), notch = TRUE) + theme_classic()
```

## ¿Is it the concentration?

```{r}
ggplot(CO2, aes(x = conc, y=uptake)) + geom_point() + theme_classic()
```

# ¿How do we determine this?

## Model formulation

```{r, echo = TRUE, eval=FALSE}
some_function(Y ~ X1 + X2 + ... + Xn, data = data.frame)
```

* **Y:** Response variable ($CO_2$ Uptake)
* **`~`:** Explained by
* **$X_n$:** Explainatory variable n (Subespecies, treatment, etc.)
* **data.frame:** data base (CO2)
* **some_function:** The model to test (our simplification of reality)

## Some models in R

```{r Modelos, echo = FALSE}
Models <- data.frame(Modelos = c("t-test" ,"ANOVA", "Linear model", "Generalized linear model", "Generalized aditive model", "non-linear model", "Mixed effect models", "Boosted regression trees"), Funcion = c("t.test()", "aov()", "lm()", "glm()", "gam()", "nls()", "lmer()", "gbm()"))

kable(Models, row.names = FALSE) %>% kable_styling(bootstrap_options = "striped")
```

## ¿Which one do we use to study the plant?

```{r, echo = TRUE}
Fit1 <- lm(uptake ~ Type, data = CO2)
```

* For this exercise we will start by using a simple linear model
* Equivalent to an ANOVA

## Using broom to get more out of your model (glance)

* To see general data of the model

```{r, eval = FALSE}
library(broom)
glance(Fit1)
```

```{r, echo=FALSE}
library(broom)
kable(glance(Fit1), digits = 2) %>% kable_styling(bootstrap_options = "striped")
```

## Using broom to get more out of your model (tidy)

* To see model parameters

```{r, eval = FALSE}
tidy(Fit1)
```

```{r, echo=FALSE}
kable(tidy(Fit1)) %>% kable_styling(bootstrap_options = "striped")
```


## Using broom to get more out of your model (augment)

* To view model predictions and residuals

```{r, eval = FALSE}
augment(Fit1)
```

```{r, echo=FALSE}
knitr::kable(augment(Fit1)) %>% kable_styling(bootstrap_options = c("striped")) %>% scroll_box(height = "300px", width = "900px")
```

## Model Selection

* Based on information criteria
* We will work with AIC
* $K$ number of parameters
* $\ln{(\hat{L})}$ fit, more positive better, more negative is bad

$$AIC = 2 K - 2 \ln{(\hat{L})}$$

## Candidate models

```{r, echo = FALSE}
ggplot(CO2, aes(x = conc, y = uptake)) + geom_point(aes(color = Type, shape = Treatment), size = 2)
```


## Candidate models

```{r, echo = TRUE}
Fit1 <- lm(uptake ~ Type, data = CO2)
Fit2 <- lm(uptake ~ Treatment, data = CO2)
Fit3 <- lm(uptake ~ conc, data = CO2)
Fit4 <- lm(uptake ~ Type + Treatment + conc, data = CO2)
Fit5 <- lm(uptake ~ Type + conc + I(log(conc)), data = CO2)
Fit6 <- lm(uptake ~ Type:Treatment + conc + I(log(conc)), data = CO2)
```

# Interpreting models

## Model 1

* uptake ~ Type

```{r echo= F, fig.height=3.5, fig.width=8}
Pred1 <- CO2
Pred1$Pred <- predict(Fit1, CO2) 
Pred1$SE <- predict(Fit1, CO2, se.fit = T)$se.fit 

ggplot(Pred1, aes(x = conc, y = Pred)) + geom_ribbon(aes(ymax = Pred + SE, ymin = Pred - SE, fill = Type), alpha = 0.5) + geom_path(aes(color = Type, )) + geom_point(aes(x = conc, y = uptake, color = Type)) + theme_bw() 
```

```{r, echo=F}
equatiomatic::extract_eq(Fit1, use_coefs = T)
```


## Model 2

* uptake ~ Treatment

```{r, echo = F, fig.height=3.5, fig.width=8}
Pred2 <- CO2
Pred2$Pred <- predict(Fit2, CO2) 
Pred2$SE <- predict(Fit2, CO2, se.fit = T)$se.fit 

ggplot(Pred2, aes(x = conc, y = Pred)) + geom_ribbon(aes(ymax = Pred + SE, ymin = Pred - SE, fill = Treatment), alpha = 0.5) + geom_path(aes(color = Treatment)) + geom_point(aes(x = conc, y = uptake, color = Treatment)) + theme_bw() 
```

```{r, echo=F}
equatiomatic::extract_eq(Fit2, use_coefs = T)
```

## Modelo 3

* uptake ~ conc

```{r, echo = F, fig.height=3.5, fig.width=8}
Pred3 <- CO2
Pred3$Pred <- predict(Fit3, CO2) 
Pred3$SE <- predict(Fit3, CO2, se.fit = T)$se.fit 

ggplot(Pred3, aes(x = conc, y = Pred)) + geom_ribbon(aes(ymax = Pred + SE, ymin = Pred - SE), alpha = 0.5) + geom_path() + geom_point(aes(x = conc, y = uptake)) + theme_bw() 
```

```{r, echo=F}
equatiomatic::extract_eq(Fit3, use_coefs = T)
```

## Modelo 4

* uptake ~ Type + Treatment + conc

```{r, echo = F, fig.height=3.5, fig.width=8}
Pred4 <- CO2
Pred4$Pred <- predict(Fit4, CO2) 
Pred4$SE <- predict(Fit4, CO2, se.fit = T)$se.fit 

ggplot(Pred4, aes(x = conc, y = Pred)) + geom_ribbon(aes(ymax = Pred + SE, ymin = Pred - SE, fill = Type), alpha = 0.5) + geom_path(aes(lty = Type)) + facet_grid(~Treatment) + geom_point(aes(x = conc, y = uptake, color = Type)) + theme_bw() 
```

```{r, echo=F}
equatiomatic::extract_eq(Fit4, use_coefs = T, wrap = T, terms_per_line = 2)
```

## Modelo 5

* uptake ~ Type + conc + I(log(conc))

```{r, echo = F, fig.height=3.5, fig.width=8}
Pred5 <- CO2
Pred5$Pred <- predict(Fit5, CO2) 
Pred5$SE <- predict(Fit5, CO2, se.fit = T)$se.fit 

ggplot(Pred5, aes(x = conc, y = Pred)) + geom_ribbon(aes(ymax = Pred + SE, ymin = Pred - SE, fill = Type), alpha = 0.5) + geom_line(aes(lty = Type)) + geom_point(aes(x = conc, y = uptake, color = Type)) + theme_bw() 
```

```{r, echo=F}
equatiomatic::extract_eq(Fit5, use_coefs = T)
```

## Model 6

* uptake ~ Type:Treatment + conc + I(log(conc))

```{r, echo = F, fig.height=3.5, fig.width=8}
Pred6 <- CO2
Pred6$Pred <- predict(Fit6, CO2) 
Pred6$SE <- predict(Fit6, CO2, se.fit = T)$se.fit 

ggplot(Pred6, aes(x = conc, y = Pred)) + geom_ribbon(aes(ymax = Pred + SE, ymin = Pred - SE, fill = interaction(Treatment, Type)), alpha = 0.5) + geom_line(aes(lty = interaction(Treatment, Type))) + geom_point(aes(x = conc, y = uptake, color = interaction(Treatment, Type))) + theme_bw() 
```

```{r, echo=F}
equatiomatic::extract_eq(Fit6, use_coefs = T, wrap = T, terms_per_line = 3)
```

# Seleccion de modelos

## Selección de modelos con broom

```{r, echo = TRUE}
Model1 <- glance(Fit1) %>% dplyr::select(r.squared, AIC) %>% mutate(Model = "Fit1")
Model2 <- glance(Fit2) %>% dplyr::select(r.squared, AIC) %>% mutate(Model = "Fit2")
Model3 <- glance(Fit3) %>% dplyr::select(r.squared, AIC) %>% mutate(Model = "Fit3")
Model4 <- glance(Fit4) %>% dplyr::select(r.squared, AIC) %>% mutate(Model = "Fit4")
Model5 <- glance(Fit5) %>% dplyr::select(r.squared, AIC) %>% mutate(Model = "Fit5")
Model6 <- glance(Fit6) %>% dplyr::select(r.squared, AIC) %>% mutate(Model = "Fit6")
Models <- bind_rows(Model1, Model2, Model3, Model4, Model5, Model6) %>% arrange(AIC) %>% mutate(DeltaAIC = AIC-min(AIC))
```

## Model selection with broom

```{r, echo = FALSE}
kable(Models)  %>% kable_styling(bootstrap_options = "striped")
```

# Any other ideas?

## Individuals

```{r, echo = F}
ggplot(CO2, aes(x = conc, y = uptake, group = Plant)) + geom_line(aes(color = Type, lty = Treatment)) + geom_point(aes(x = conc, y = uptake, color = Type, shape = Treatment)) + theme_bw() 
```


## Mixed effect models

```{r}
library(lme4)

Fit7 <- lmer(uptake ~ Type:Treatment + conc + I(log(conc)) + (1 | Plant), CO2)

```


```{r, echo = FALSE}
broomExtra::glance_performance(Fit7) %>% dplyr::select(-sigma, -bic, -remlcrit) %>% kable() %>% kable_styling(bootstrap_options = "striped")
```

