---
title: Using Spatiotemporal GLMMs through Gaussian Markov Random Fields and Meshes to Enhance Spatially Explicit Bayesian Modeling
author: "Derek Corcoran"
date: "`r Sys.Date()`"
output:
  ioslides_presentation:
    widescreen: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE,
                      error = FALSE,
                      message = FALSE,
                      cache = T,
                      tidy = 'formatR')
library(SpatioTemporalCont)
library(terra)
library(tidyterra)
library(ggplot2)
library(purrr)
library(sdmTMB)
library(sdmTMBextra)
library(patchwork)
library(kableExtra)
library(DHARMa)
library(readr)
```

# Problem

## Spatially explicit models

* Spatially explicit models are often not spatially explicit
* We often just transform it to tabular data
* We have to go out of our wa 


## Spatially explicit GLMM

```{r mesh}
data("pcod")
bnd <- INLA::inla.nonconvex.hull(cbind(pcod$X, pcod$Y), convex = -0.1)
mesh_inla <- INLA::inla.mesh.2d(
  boundary = bnd,
  max.edge = c(25, 50)
)
mesh <- make_mesh(pcod, c("X", "Y"), mesh = mesh_inla)
plot(mesh)
```

## First model

```{r}
m <- sdmTMB(
  data = pcod,
  formula = present ~ depth_scaled + + I(depth_scaled^2),
  mesh = mesh, # can be omitted for a non-spatial model
  family = binomial(link = "logit"),
  spatial = "off"
)
```


## Spatial residuals

```{r SpatialResiduals}
coords <- pcod[,c("X", "Y")]
coords$Resid <- residuals(m)
```


```{r SpatialResidualsplot, echo=FALSE}
G1 <- ggplot(data = coords, aes(X, Y)) + geom_point(aes(color = Resid))  + scale_color_gradient2(name = "residuals", limits = c(-4,4)) + labs(title = "Spatial off")  +
  coord_fixed() + theme(legend.position = "none")

G1
```

## Spatial on

```{r}
m1 <- sdmTMB(
  data = pcod,
  formula = present ~ depth_scaled + + I(depth_scaled^2),
  mesh = mesh,
  family = binomial(link = "logit"),
  spatial = "on"
)
```


## Spatial residuals

```{r SpatialResiduals2}
coords$ResidSpatial <- residuals(m1)
```


```{r SpatialResidualsplot2, echo=FALSE}
LongCoords <- coords |> 
  tidyr::pivot_longer(Resid:ResidSpatial, values_to = "Residual") |> 
  dplyr::mutate(Spatial = ifelse(name  == "Resid", "Spatial: off", "Spatial: on"))

G2 <- ggplot(data = LongCoords, aes(X, Y)) + geom_point(aes(color = Residual))  + scale_color_gradient2() +
  coord_fixed() + facet_wrap(~Spatial)

plotly::ggplotly(G2)
```


## Lets look at the parameters

```{r tidies}
t <- tidy(m,  conf.int = T) |> 
  mutate(Spatial = "off")
t1 <-tidy(m1,  conf.int = T) |> 
  mutate(Spatial = "on")

tidies <- dplyr::bind_rows(t, t1)
```

```{r tableTidies1, echo=FALSE}
kbl(tidies) |> kableExtra::kable_paper()  |> scroll_box()
```

## Add Spatiotemporal random fields

```{r spatiotemporal}
m2 <- sdmTMB(
  data = pcod,
  formula = present ~ depth_scaled + I(depth_scaled^2),
  mesh = mesh,
  family = binomial(link = "logit"),
  spatial = "on",
  time = "year",
  # first-order autoregressive 
  spatiotemporal = "AR1",
  # Show that there are missing years
  extra_time = c(2006, 2008, 2010, 2012, 2014, 2016)
)
```

## Check for issues {.smaller}

```{r sanity1, message=TRUE, warning=T, error=TRUE}
sanity(m2)
```

## Calculate response


```{r prediction}
NewDf <- data.frame(depth_scaled = seq(from = min(pcod$depth_scaled), to = max(pcod$depth_scaled), length.out = 100), year = 2015)

p <- predict(m2, newdata = NewDf, se_fit = TRUE, re_form = NA, type = "response")

p <- p |> mutate(prob = 1 / (1 + exp(-est)), upper = 1 / (1 + exp(-(est + 1.96*est_se))), lower = 1 / (1 + exp(-(est - 1.96*est_se))))
```

## Plot response


```{r plotResponse, echo = F}
ggplot(p, aes(x = depth_scaled, y = prob)) + geom_ribbon(aes(ymax = upper, ymin = lower), fill = "red", alpha = 0.5) + geom_path() + theme_bw()
```

## Time varying response


```{r timevaryingMod}
m3 <- sdmTMB(
    data = pcod,
    formula = present ~ 0 + as.factor(year),
    time_varying = ~ 0 + depth_scaled + I(depth_scaled^2),
    mesh = mesh,
    family = binomial(link = "logit"),
    spatial = "on",
    time = "year",
    # first-order autoregressive 
    spatiotemporal = "AR1",
    # Show that there are missing years
    extra_time = c(2006, 2008, 2010, 2012, 2014, 2016)
)
```


## Prediction time varying


```{r predictionVarying}
NewDf <- expand.grid(depth_scaled = seq(from = min(pcod$depth_scaled), to = max(pcod$depth_scaled), length.out = 50), year = unique(pcod$year))

p <- predict(m3, newdata = NewDf, se_fit = TRUE, re_form = NA, type = "response")

p <- p |> mutate(prob = 1 / (1 + exp(-est)), upper = 1 / (1 + exp(-(est + 1.96*est_se))), lower = 1 / (1 + exp(-(est - 1.96*est_se))))
```


## Plot time varying response

```{r plotResponseVarying, echo = F}
ggplot(p, aes(x = depth_scaled, y = prob)) + geom_ribbon(aes(ymax = upper, ymin = lower), fill = "red", alpha = 0.5) + geom_path() + theme_bw() + facet_wrap(~year)

```

## Quick model comparison

```{r ModelSel, eval =F}
Sel <- MuMIn::model.sel(list(m, m1, m2,m3)) |> 
  as.data.frame() |> 
  dplyr::select("spatial", "time", "spatiotemporal", 
"time_varying", "df", "AICc", "delta", "weight")
```

```{r ModelSel2, eval =T, include=FALSE}
Sel <- MuMIn::model.sel(list(m, m1, m2,m3)) |> 
  as.data.frame() |> 
  dplyr::select("spatial", "time", "spatiotemporal", 
"time_varying", "df", "AICc", "delta", "weight") |> 
  dplyr::mutate(time_varying = as.character(time_varying),
                time_varying = ifelse(time_varying == "", "No", "Yes"))
```
```{r tableModSel, echo=FALSE, cache = F}
kbl(Sel, digits = 2) |> kableExtra::kable_paper()  |> scroll_box()
```
